How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-09-05 1.00-Part1.Rmd
Add files via upload
#general update of all files

@@

# (PART) Part I: Getting ready for submission {-}

# Submitting your work to a journal {- #part1}

Part 1 of this book is about the publication arena. I provide simple explanations of what scientific journals are for, and who the gatekeepers are. The publishing world is at a turning point, and before you start publishing you should be aware of the current trends and what is possible.

>>>>

# (PART) Part 1: Right from the very start {-}

@@

# Introduction  {#part1}

Right from the very start of your PhD, it is important for you to become familiar with the goal of what you are trying to achieve, how you plan to achieve it, and (believe it or not) what you plan to do afterwards. There are a lot of things to think about as you start your PhD, and for many of these it is far easier to make informed decisions from the outset. For example, reconciling yourself to using the [statistical package R](#useR) will be much easier from the start, rather than only realising under pressure at the end of your PhD that there is no way that you can finish unless you take it on. Preparing yourself into the right mental framework will also be very important for your ability to work under pressure. Getting into good habits at the beginning will put you in a commanding position later. The aim of this first part of the book is to help you assess your current position, and realign yourself toward an optimum mindset.

### 2021-09-05 1.01-beforeyoustart.Rmd
Add files via upload
#general update of all files

---
output:
  html_document: default
  pdf_document: default
---

\mainmatter

# Introduction  {#intro}

Right from the very start of your PhD, it is important for you to become familiar with the goal of what you are trying to achieve, how you plan to achieve it, and (believe it or not) what you plan to do afterwards. There are a lot of things to think about as you start your PhD, and for many of these it is far easier to make informed decisions from the outset. For example, reconciling yourself to using the [statistical package R](#useR) will be much easier from the start, rather than only realising under pressure at the end of your PhD that there is no way that you can finish unless you take it on. Preparing yourself into the right mental framework will also be very important for your ability to work under pressure. Getting into good habits at the beginning will put you in a commanding position later. The aim of this first part of the book is to help you assess your current position, and realign yourself toward an optimum mindset.

## So you are doing a PhD? {#beforeyoustart}

>>>>

# So you are doing a PhD? {#beforeyoustart}

### 2021-09-05 2.00-Part2.Rmd
Add files via upload
#general update of all files

@@

## How to write a hypothesis  {#hypothesis}

When you read about the formula of writing a chapter (see [end of part 2](#formula)) you will see that everything in your chapter or paper revolves around the hypothesis or questions that you ask, normally at the end of your introduction (Figure \@ref(fig:Hypothesis)). In this section, we will take a look at how to write a hypothesis. This is a sticking point for many students. We are used to using and writing questions and statements in day-to-day communications, as well as reading popular media. But hypotheses (the plural of hypothesis) only rarely float across our desks. So how do we write one, and how do we know if our hypothesis is good?

In addition to this section, there is some good information out there on the web, and it's worth looking at this too: (e.g. [Wikihow](https://www.wikihow.com/Write-a-Hypothesis), [Wikipedia](https://en.wikipedia.org/wiki/Hypothesis), etc.). There's also some less good stuff out there, so read critically.

(ref:Hypothesis) **Generating a good hypothesis for your PhD chapter isn't easy.** A good hypothesis will be invaluable in helping you write the chapter. It is important to start working with hypotheses.

```{r Hypothesis, echo=FALSE, fig.align='center', out.width='90%', fig.cap="(ref:Hypothesis)"}
knitr::include_graphics('figures/hypothesis.jpg')
```

### What is a hypothesis?  {#whathypothesis}

A hypothesis is a statement of your research intent. It tells the reader (because just like all of your other written work, it is intended for an audience who will read it), what you plan to do in your research. But there's a little more to it than this. The hypothesis becomes a part of the scientific method if it is testable and (importantly) falsifiable, as well as being informed from previously published work on the subject.

Your hypothesis must be informed by the literature, which is why you spent so much time and effort crafting your introduction to inform your reader of the same. This is also why your hypothesis usually comes at the end of your introduction, because you spend all of the introduction telling your reader about it. There's not much point in writing more after the hypothesis, because once your reader has read that, they are ready to learn about how you went about testing it (in the [Materials & Methods](#methods)). The other important point to make is that the literature should dictate how you write your hypothesis, and the variables that you include. If, for example, you know that temperature is the most important variable, but all of the literature suggests that it is oxygen, you can't ignore oxygen and you should also frame your hypothesis using this variable (you can have more than one hypothesis after all!). In this case, you will also need to provide a sufficient introduction to temperature as a variable to justify its inclusion in your hypothesis. Perversely, your aim is not to prove that your idea is right, but to show that the hypothesis can be refuted.

We try to write a hypothesis that is falsifiable: i.e. you can prove (usually using statistical tests) that it is not correct (or at least show that the likelihood that it is correct is very low). That's why it is conventional to provide the 'null hypothesis' that is the falsified version of the statement, suggesting that there is no relationship between the variables you have proposed to measure. The convention is to label this 'null hypothesis' H~0~, while the 'alternative hypothesis' (the one that says your variables are related as you suggested) is written as H~1~. When you formulate your hypothesis, it is traditional to write your alternative hypothesis to indicate the directionality of your tested variables. This way, the reader can simply imply that the null hypothesis is when there is no relationship, but this will need to be stated if the null hypothesis is more complex.

Karl Popper [-@popper2005logic] was the philosopher who proposed that without being able to refute or falsify a scientific problem, it ceases to be scientific. This is the reason for our null hypothesis. If the null is not available as a possible outcome, then logically, there is no science.

> Karl Popper [-@popper2005logic]:  
> **"...it must be possible for an empirical scientific system to be refuted by experience". **

Most importantly, your hypothesis must come first, before you do the experiment or study. Hence the reason why this section comes at the start of part 2. Setting the hypothesis after the work is already done is fraudulent, and goes against the scientific method. Obviously, it isn't fair to pose the hypothesis once you already know the answer (also known as [HARKing](#TypeI)). This is why there is so much emphasis put on formulating your hypothesis during your research proposal. Getting it right will determine what you do and how you test it. If you think of an extra hypothesis that would be really useful to test once you've already done your study, you can conduct a post hoc test, but this should have more stringent levels of statistical assessment.

Writing a hypothesis isn't easy, but it is essential, and once you've understood what to do most of the rest of what you are writing for should make sense.

### What a hypothesis isn't

It is not a question and so should never have a question mark after it.

It isn't really a simple prediction: if this then that. You will see on the internet that hypotheses are explained in this simple predictive framework. I say that a hypothesis is not a simple prediction because it lacks the mechanistic and scholarly aspect of a good hypothesis, which is what we want to achieve.

### A formulaic way to start writing your hypothesis:

__"If. then. because."__

Later in the book, I will emphasise that you must have introduced all the variables that you plan to use to test your hypothesis in your introduction. This usually comes in the second paragraph of your introduction, where you emphasise the utility of the dependent variable/s (what you are planning to measure) and your independent variable (what you will manipulate). Both of these variables should then feature in your hypothesis. Next, by paragraph four, you will have identified the problem that you are interested in tackling. In addition, your introduction will provide all of the pertinent literature that has relevance to this hypothesis, giving the all important context.

A simple way to consider making your hypothesis is to adopt an **"If. then. because."** construction where you add in your problem statement using your independent variable after 'if', and your prediction using your dependent variable after 'then', and finally the expected mechanism after 'because'. Using our example above with the "If. then. because." construction, we would say: "If environmental temperatures in which tadpoles develop are increased then tadpole development rate is faster because they follow the classic metabolism of ectotherms". Both independent variable (temperature) and dependent variable (tadpole development rate) are present in this hypothesis, and the predicted relationship between them is clear. In addition, the causal mechanism is stated. I say that this is a formulaic way to start writing your hypothesis, because it usually ends up as an inelegant statement, which can be better refined for a reader. A citation for your stated mechanism might also help clarify exactly where the justification for this comes from.

Mechanisms (or causal explanations) fall into three main areas: endogenous, exogenous and evolutionary [@allen2017scientific].

#### Endogenous causal explanations

Endogenous causal explanations focus on the mechanisms happening inside an organism, such as physiological processes, hormones, reproductive state, etc.

#### Exogenous causal explanations

Exogenous causal explanations concern mechanisms that are outside the body of individuals. Common exogenous mechanisms are climatic factors (temperature, humidity, precipitation, etc.) or may relate to the availability of food, predators or mates.

#### Evolutionary causal explanations

These mechanisms have evolved through time, and often relate to exogenous mechanisms triggering endogenous processes over multiple generations.

Note then that the above mechanisms are not mutually exclusive in their nature, and it may be useful to combine different approaches within biology to ask hypotheses across all of these levels. Mechanisms in biological sciences are rarely simple or act on multiple organismal levels, so designing a controlled experiment in order to test a specific mechanism thoroughly can be very demanding. In other words, can you be sure that the cause is really responsible for the effect that you are measuring?

A good hypothesis will often take an existing hypothesis further, to try to better refine the knowledge on a subject. Hence, it is perfectly acceptable to state that you are building on existing hypotheses (and giving the appropriate statement) when making your own.

### Teleological versus causal hypotheses

A teleological argument refers to the reason or a purpose of a particular process. For example, you may measure vertical migration of water fleas and suggest that diurnal migrations are made because the water fleas want to avoid predation. This is a teleological hypothesis because you are suggesting that the reason behind a process is the desire by water fleas to avoid predators. Although a reduction of predation may be a consequence of vertical migration in water fleas, each water flea does not think about predation and then starts it’s upward movement as a result. A common mistake made in biology is to apply teleological arguments to processes that have no purpose or reason. Evolution is often mistakenly suggested to have a purpose (e.g. to evolve to a more advanced state), but in fact, evolution is not a goal-orientated process. There is no end-point to evolution, and evolution did not start in order to meet some predetermined form or function. On the other hand, a causal hypothesis focuses on the factors about A that cause B.

You should have realised that biologists are principally interested in causal hypotheses, because most mechanisms that are studied in biological sciences have no predetermined goal. If you are a behavioural ecologist, then you will need to be particularly aware of these two types of hypotheses, and when teleological explanations may be appropriate: many types of behaviour are goal orientated.

### How to evaluate your hypothesis

Once you've written your hypothesis, how do you decide whether or not it is good? To do this, you might think that you need plenty of experience (and yes, that does help). But really, you just need to look for the elements that are discussed above. So once you've written your hypothesis, try to objectively answer the questions below (for more see [here](https://youtu.be/9KoSGQlIFkM)):

- Is there a clear prediction (if. then. statement)?
- Does the prediction use independent and dependent variables correctly?
- Is the mechanism supported by the literature?
- Is the hypothesis testable/falsifiable?
- Does the hypothesis use concise wording and precise terminology?

If your hypothesis meets all of the criteria above, then you've done a good job!

Probably one of the hardest issues that you will face in biological sciences is to determine whether your dependent variable is reliant only on your independent variable of choice. For example, a lot of variance in biological sciences will relate to the climate (especially with global change studies), but if your independent variable is temperature, this means that you will need to keep all other climatic variables the same. That is, if temperature is your independent variable, it is the only variable that can change in your experiment. This type of experiment is challenging as temperature often affects other variables (especially that they may vary in an unpredictable way). As soon as you have more than one independent variables, you can no longer test your dependent variable because you don't know which independent variable it is reacting to.  Isolating variables is notoriously difficult, especially when we move from the laboratory to the field. You will need to think very carefully about what variables other than those of interest are potentially impacted by your experimental design. If you cannot control for them, this will likely mean that you need to change your hypothesis, or change your experimental design. If you are unsure, then I would encourage you to look carefully at the experiments that others have conducted in the literature.

>>>>

# (PART) Part 2: Get writing {-}

@@

### 2021-09-05 2.01-Hyp.Rmd
Add files via upload
#general update of all files

# How to get started with writing  {#part2}

__Before you start writing__

The writing part of the PhD is often the most difficult for many students. I think that there are several reasons for this. Writing well comes with practice, and most PhD students have had little experience with academic writing even if they have already completed an MSc. The same could be said of most early-career scientists (even after their PhDs), because practice is measured in completed pieces, it usually takes many years rather than weeks or months. You shouldn't expect that scientific writing will come automatically or immediately. In order to get this practice, you need to get started, and (again) many students find starting to write challenging. This guide was written to help you get started.

However, before you start writing, it's a good idea to understand some of the ways in which academic writing differs from other writing styles. Good planning is the key here, along with knowing exactly what it is that you'll be writing about. That sounds facile, but at the foundation of each scientific study lies the hypothesis, which is central to being able to build the rest of the chapter.

In this part of the book, I start with an explanation of the hypothesis. As you will see, the hypothesis is central to all of the writing in each chapter or paper, and you need to have a very good idea of how to formulate your research question into a hypothesis before you begin. Then I build on helping you think through lots of different aspects of writing in science: style, citations, writing concisely, scientific names, and lots more. Finally, I introduce you to the idea of using a formula to structure your writing.

Before you start writing in earnest, it’d be good to read this part of the book and make sure that you are familiar with the ideas herein. Later, you may need to refer back to sections once you’ve started writing.

## How to write a hypothesis  {#hypothesis}

>>>>

# How to write a hypothesis  {#hypothesis}

### 2021-09-05 2.12-taxonomy.Rmd
Add files via upload
# corrections after July 2021 Copy Edits
Add files via upload
#general update of all files

It might seem that there's a mistake above, because Tschudi, 1838 isn't in brackets. This is actually deliberate. Tschudi, 1838 isn't in brackets because the genus name hasn't changed since the description by Tschudi [1838](tschudi1838classification) placed *S. capensis* in *Sclerophrys* in 1838. But it is clear in the table above that the genus name did change to *Amietophrynus*, and it was recognised under a different species name (*A. rangeri*). Unlikely as it may seem, the original description of the species was by Tschudi in 1838, but this was long forgotten (or ignored because the type specimen was so poorly preserved), and the species was regularly referred to as *Bufo rangeri*. When the genus *Bufo* was broken up into lots of newly named genera by Frost et al [-@frost2006amphibian], Tschudi's description was once again disregarded [@ohler2016identity]. However, in 2016 new work on the type specimen revealed that it was the same as *Amietophrynus rangeri,* but because Tschudi's description was older, the taxonomy deferred to the rule of precedence. If you want to read more about it, you can look at all the changes on the amphibian taxonomic database.

@@

### View 1

Common names are proper nouns and proper nouns start with capitals. The Western Leopard Toad is not any toad, and the African Clawed Frog isn't any frog. The Cape platanna is any platanna living in the Cape, but the Cape Platanna is *Xenopus gilli*! In this last example, you can see how capitalisation reduces ambiguity and differentiates between adjectives and nouns.

### View 2

Common names are not formal names (like scientific names) so why does it matter? Many journals do not capitalise common names and insist on lower case unless they are proper nouns, like African in African clawed frogs. This seems to have arisen from editors who dislike superfluous capitalisation.

@@

>>>>

It might seem that there's a mistake above, because Tschudi, 1838 isn't in brackets. This is actually deliberate. Tschudi, 1838 isn't in brackets because the genus name hasn't changed since the description by Tschudi [-@tschudi1838classification] placed *S. capensis* in *Sclerophrys* in 1838. But it is clear in the table above that the genus name did change to *Amietophrynus*, and it was recognised under a different species name (*A. rangeri*). Unlikely as it may seem, the original description of the species was by Tschudi in 1838, but this was long forgotten (or ignored because the type specimen was so poorly preserved), and the species was regularly referred to as *Bufo rangeri*. When the genus *Bufo* was broken up into lots of newly named genera by Frost et al [-@frost2006amphibian], Tschudi's description was once again disregarded [@ohler2016identity]. However, in 2016 new work on the type specimen revealed that it was the same as *Amietophrynus rangeri,* but because Tschudi's description was older, the taxonomy deferred to the rule of precedence. If you want to read more about it, you can look at all the changes on the amphibian taxonomic database.

@@

### View 1 - Yes, use capitals for common names

Common names are proper nouns and proper nouns start with capitals. The Western Leopard Toad is not any toad, and the African Clawed Frog isn't any frog. The Cape platanna is any platanna living in the Cape, but the Cape Platanna is *Xenopus gilli*. In this last example, you can see how capitalisation reduces ambiguity and differentiates between adjectives and nouns.

Personally, I find that capitalising common names helps them stand out in the text, and as a reader it helps me immediately recognise that a word is a name, and not a regular adjective. As many common names contain strings of words including adjectives, I find that I might need to re-read a section when I realise that a common name is being used, and that the author is not describing an attribute of the species.

Note that many common names are hyphenated, and if this is the case then you should only capitalise the first part of two in a hyphenation: i.e. **White-throated Sparrow** (_Zonotrichia albicollis_), not White-Throated Sparrow.

### View 2 - No, don't use capitals for common names

Common names are not proper nouns (like scientific names) and so should not be capitalised. Many journals do not capitalise common names and insist on lower case unless they are proper nouns, like African in African clawed frogs. This seems to have arisen from editors who dislike superfluous capitalisation.

Other reasons why capitalisation might not be implemented for common names include:

- They are not proper nouns, and so (according to traditional English grammar rules) should not be capitalised
  - Countries publish style manuals for what is acceptable (e.g. see [here](https://www.stylemanual.gov.au/)), and these often state that common names should not be capitalised
  - I suspect that many publishers follow these official national guidelines when making decisions about scientific journals
  - Many publishers and their editors hate superfluous capitalisation and attempt to eliminate it. This simply comes from sloppy writers who tend to overcapitalise. 
- The capitalisation of common names implies some sort of official recognition: that it is **the** common name, and not **any** common name. Some groups, such as ornithologists, have only one common name for each species that they officially recognise. They take this to absurd lengths when they insist that all common bird names **must** be capitalised while common names for other taxonomic groups must not be capitalised (see [here](https://www.worldbirdnames.org/new/english-names/spelling-rules/capitalization/)).
  - Even when ornithologists get their own way, and bird names get capital letters, these are changed to lower case when they are in a list with other animals and plants that carry lower-case (see [here](https://www.stylemanual.gov.au/style-rules-and-conventions/names-and-terms/plants-and-animals#write_common_names_in_lower_case_and_roman_type))
- You may notice that Wikipedia does not use capitalisation for common names (see [here](https://en.wikipedia.org/wiki/Wikipedia:Manual_of_Style/Capital_letters#Common_names)). They state that this is because they avoid superfluous capitalisation in English (much as editors do - see above).

@@

### Changing scientific names to English names

There are many examples of names of scientific taxa that are changed to English derivatives of the original Latin names (see Table \@ref(tab:derivative)). While the official Latin version of the scientific name almost always carries a capital letter (see [above](#taxonomy)), these anglicised versions do not.

Table: (\#tab:derivative) **Latin names can be changed into English versions.** You may not realise it, but many English names have been derived from scientific names. While the Latin names usually carry capital letters (see [above](#taxonomy)), these anglicised versions do not.

| Taxonomic rank | Latin name | English derivative |
| ---- | -- | --- |
| Kingdom | Animalia | animal |
| Phylum | Arthropoda | arthropod |
| Order | Carnivora | carnivore |
| Superfamily | Vespoidea | vespoids |
| Family | Bufonidae | bufonid |
| Subfamily | Felinae | feline |
| Genus | _Eucalyptus_ | eucalypt |

Turning a Latin taxonomic rank into an English derivative is a useful mechanism in which to bring variation into your writing. While there are distinct rules about the use of taxonomic ranks (see [above](#taxonomy)), there are not with these anglicised derivatives, and you can use them as you want. Even if they haven't been used before for your study group, there is no reason why you should not erect an English derivative for your own purposes - as long as this enhances readability and does not cause more confusion.

### 2021-09-05 2.17-plagiarise.Rmd
Add files via upload
#general update of all files

>>>>

## Free online plagiarism tools

If your institution doesn't subscribe to a plagiarism tool, then you can check your work online free using one of a number of tools available. There are advantages and disadvantages to each, so I suggest that you try them out, and see which suits your needs best.

- **Grammerly** ([www.grammarly.com](www.grammarly.com)) Checks your work for plagiarism for free, but you need to create an account (even for the free version) and you have to pay if you want it to check your grammar. 
- **Plagiarisma** ([plagiarisma.net](plagiarisma.net)) You'll need to create an account (to see where your copied text came from) and paste your text into a box, but you'll get a good sense of whether or not there is plagiarism in your text.
- **PaperRater** ([www.paperrater.com](www.paperrater.com)) No need to create an account to paste text into a box (but it didn't recognise my content which is also on my blog and so doesn't get top marks from me). Some nice different output with suggestions on readability and grammar.

### 2021-09-05 3.00-Part3.Rmd
Add files via upload
#general update of all files

@@

## Title page {#title}

Most of the following information in this section on the title pertain to getting a manuscript ready for submission to a journal. For your thesis, I would recommend that you have a title page for each chapter that lists any collaborators and their contributions so that this is transparent for the examiners. It also helps break up the document and can be a good spot for a non-academic illustration (if your institution permits).

Nearly all journals will require you to have a title page for your manuscript. This may or may not include authors. Check to see whether the journal that you are submitting to conducts double-blind review. If they do, they will require a title page without any indication of authors or institutions (they will also likely ask you to remove the acknowledgements section). This should be clearly specified in the instructions to authors of your chosen journal.

### Names and addresses are important {#ORCID}

Getting authors' names correct, their correct addresses (many people have more than one affiliation), can be tricky. Check that you have all of the information required for a title page before you submit. While you need the names of your co-authors to be displayed correctly, you do not need their titles (no Dr. or Prof., etc.).

Indicate clearly who is the corresponding author. Normally, if you have done all of the rest of the work, the corresponding author should be you. You need to learn how to start this role at some point, so it might as well be now. Having the correct name and address will be important to each author. Along with the author names and addresses you should record the ORCID of each author. The ORCID is simply a unique identification code for individual researchers. ORCID is a non-profit organisation and there’s nothing sinister in signing up. If they (or you) don’t have one, then you should ask them to create one before the submission. It takes less than 5 minutes. Go to www.orcid.org

## The title

The title of your paper, chapter or book is the first thing that any reader will read, and so should be well considered. If your reader cannot quickly understand your title, the chances are that they won’t bother reading any further. Your title will be your selling point, and your aim is to use it to draw your readership in. Once you’ve managed to inveigle your readers to download your paper, your title is also their hook for remembering your paper in their database of thousands of others. Having their key-word in your title will help here, and as ever with writing your challenge is to think like your reader. The best titles are those that sum up the entire study in five to seven words. This is best done in a narrative that tells the story ([see part 2](#storytelling)) of your manuscript in its entirety. This may sound daunting, but you should get into the habit of summing up your story quickly (for friends, relatives and work colleagues). Then it’s a question of refining this story into the short single sentence that makes up the title. While the narrative approach may not work for you, you do want the title to provide enough information so that the potential reader knows what they will find before they open it. Your title doesn’t just have to work for you, it needs to work for a wide audience.

Some people are excellent at writing titles that contain puns of well-known phrases or sayings. These can be brilliant, working both to inform what’s in the paper as well as providing some familiar input that helps retain them in memory. However, many fail to do either and are simply a waste of space. If you are tempted to use a pun as your title, make sure that it is widely appreciated, and not just among your co-authors and lab.

You don’t have to come up with a killer title from day one. Most of my manuscripts have a working title that gets revised as I write, and is always open for change before submission. If you have great ideas for a title, do note them down. I find that the more options I have, the more likely I am to come up with something that works for everyone. It also helps to mix and match from a set of candidate titles. Once you have come up with something that looks good to you and your colleagues, test it by entering it into your database of choice (with the default being Google Scholar). Your first 10 results should include a set of papers that you have likely cited in the upper area of your introduction. If you don’t recognise any out of the top 10, it’s time to look at another of your candidate titles.

### Some title ideas to start you off
- Don’t start by looking for the best title
  + Write a number of candidate titles and ask your co-authors to vote for their favourites.
- The shorter and catchier your title can be the better: 6 words (±1) is an ideal.
  + Allow yourself a longer subtitle if needed, but don’t go over 20 words total (some journals may limit your total to less). 
  + Consider the (former) 120 character limit of a Tweet as an upper limit
- Do include your principal finding if possible
- Include as many key-words as you can

### Things to avoid in your title
 - Don’t feel obliged to include [taxonomic terms](#taxonomy) unless it is relevant or compulsory
   + Some (taxonomically minded) journals will insist on the species name followed by the taxonomic authority, and/or the family and order, in your title
 - Avoid obscure or specialist words that won’t be understood by your readership
   + There are times when key words are necessarily specialised and your readership will expect this, but simpler words in your title will open up your readership which will otherwise remain narrow
 - Don’t simply define the scope of your work without including your content.
 - Don’t have your title as a question. Rather provide the answer!

## Key-words
The key-words are a way for readers to find your content with searches. Typically, the advice is to use words that are not in the title or abstract. This is because many databases have combined searching facilities for title, abstract and key-words. I struggle to think of appropriate key-words, and so make a list of some of the [big idea](#bigidea) words and short phrases from the introduction. Then I tend to look at articles in the same genre, and see what key-words they have used. As with the title, I suggest that you enter your chosen key-words into your literature database of choice [see part 2](#database) and see what comes back. You should see a group of papers that look wholly familiar and preferably those that are already cited in your manuscript. If not, it’s time to look again.

>>>>

# (PART) Part 3: The formula of a chapter {-}

@@
